
---
title: "SBC on the GBM dataset from the whole Firebrowse Database(TCGA)."
subtitle: "Autoencoder1 Molecular+Survival data"
author: "Camila Duitama (Based on Ashar Ahmad's code for the SBC model)"
date: "31 July 2019"
params:
  iter: 100
  iter.burnin: 150
  iter.thin: 5
  pval.sbc : 0.6
  output_folder : "Autoencoder1_Molecular+Survival_files/"

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE,eval=FALSE,include=TRUE,collapse=TRUE,warning=FALSE,message=FALSE, fig.width = 10, fig.height = 6)
rm(list = ls(all.names = TRUE))
#load("/Volumes/GoogleDrive/My Drive/Documents/GitHub/MasterThesis/R_Scripts_First_Results/Autoencoder1_Molecular+Survival.RData")

```
# Description

A simple autoencoder with one layer for the encoder and one layer for the decoder was built. The input of the autoencoder were both the molecular and survival data sets. One autoencoder was built for the training set, and another one for training set.This code can be found in [here](https://github.com/CamilaDuitama/MasterThesis/blob/master/Data%20preparation-RTCGA.ipynb) 


# Loading of libraries

Libraries needed to run the code must be loaded first

```{r LibrariesLoading, eval=FALSE}
library(gsva)
library(sva)
library(readr)
library(ggplot2)
library(matrixcalc)
library(stats)
library(Runuran)
library(truncnorm)
library(Matrix)
library(psych)
library(VGAM)
library(MixSim)
library(statmod)
library(flexclust)
library(mixAK)
library(mclust)
library(monomvn)
library(cluster)
library(flexmix)
library(survival)
library(utils)
library(rms)
library(pec)
library(ipred)
library(verification)
library(Hmisc)
library(glmpath)
library(glmnet)
library(gplots)
library(sparcl)
library(kernlab)
library(class)
library(reshape)
library(GGally)
library(XML)
library(RCurl)
library(MASS)
library(matrixcalc)
library(stats)
library(Runuran)
library(MCMCpack)
library(VGAM)
library(statmod)
library(survcomp)
library(gtools)
library(ggplot2)
library(GenOrd)
library(plyr)
library(dplyr)
library(iCluster)
library(CCA)
library(caret)
library(mcclust)
library(Biobase)
library(mixtools)
library(knitr)
library(kableExtra)
library(devtools)
library(rmarkdown)


```

# Loading/Splitting training and testing sets

Data is separated into training and testing

```{r Loading files, eval=FALSE}
######## Input ###############

Clinical_TrainingSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Second_data_split/TrainingSet/Clinical_TrainingSet.csv")
mRNAArray_TrainingSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Second_data_split/Autoencoders/TrainingSet/mRNAArray+survival_TrainingSet_AE1.csv")
Clinical_TestSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Second_data_split/TestSet/Clinical_TestSet.csv")
mRNAArray_TestSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Second_data_split/Autoencoders/TestSet/mRNAArray+survival_TestSet_AE1.csv")
indexes.train<-which(is.na(Clinical_TrainingSet$days_to_death), arr.ind=TRUE)
indexes.test<-which(is.na(Clinical_TestSet$days_to_death), arr.ind=TRUE)
training_time<-Clinical_TrainingSet$days_to_death
test_time<-Clinical_TestSet$days_to_death
for (i in indexes.train){
  training_time[i]<-Clinical_TrainingSet$days_to_last_followup[i]
}
for (i in indexes.test){
  test_time[i]<-Clinical_TestSet$days_to_last_followup[i]
}
total_time<-append(training_time,test_time)

n<-length(Clinical_TrainingSet$X1)
n.new<-length(Clinical_TestSet$X1)


#### Overall Survival Time Vector (N*1)
time<-total_time[1:n]
time<-log(time)
########### Event or Not Vector (N * 1) ####
censoring<-Clinical_TrainingSet$vital_status
######### mRNA (or miRNA expression values) (N*D) ###
mRNAArray_TrainingSet$X1<-NULL
Y.pre.train<-mRNAArray_TrainingSet

#### Overall Survival Time Vector (N*1)
time.new<-total_time[-(1:n)]
time.new<-log(time.new)
########### Event or Not Vector (N * 1) ####
censoring.new<-Clinical_TestSet$vital_status
######### mRNA (or miRNA expression values) (N*D) ###
mRNAArray_TestSet$X1<-NULL
Y.pre.test<-mRNAArray_TestSet

```

```{r Number of samples}
print(paste0("Number of patients in training set: ",as.character(n)))
print(paste0("Number of patients in test set: ",as.character(n.new)))
```

##Data visualization

### Censoring information
```{r Table for censoring information of the training set}

table(censoring) %>%  
  kable(caption = "Censoring frequency for the training set") %>%
  kable_styling(bootstrap_options = c("striped", "hover","responsive"),full_width = T, position = "center")

```

```{r Table for censoring information of the testing set}
table(censoring.new) %>%
  kable(caption = "Censoring frequency for the testing set") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")

```

### Header of mRNA matrixes

```{r Header of mRNA matrix of the training set}
  Y.pre.train[1:5,1:5] %>%
  kable(caption = "mRNA MicroArray data for the Training set " ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")
```

```{r Header of mRNA matrix of the testing set}
  Y.pre.test[1:5,1:5] %>%
  kable(caption = "mRNA MicroArray data for the Testing set " ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = F, position = "center")
```


### PCA of training and test sets
```{r PCA of the whole mRNA dataset}
mRNAArray_TestSet.scaled<-scale(mRNAArray_TestSet,scale=TRUE,center=TRUE)
mRNAArray_TrainingSet.scaled<-scale(mRNAArray_TrainingSet,scale=TRUE,center=TRUE)

totalmRNA<-rbind(mRNAArray_TrainingSet.scaled,mRNAArray_TestSet.scaled)
mRNA.pca<-as.data.frame(prcomp(totalmRNA,scale=TRUE, center= TRUE)$x[,1:2])
classes<-c()
for (i in 1:length(Clinical_TrainingSet$days_to_death)){
    classes<-append(classes,"Training")
  }
for (i in 1:length(Clinical_TestSet$days_to_death)){
    classes<-append(classes,"Testing")
  }

mRNA.pca$Classes<-classes
p1 <- ggplot(mRNA.pca, aes(x=PC1, y=PC2, colour=Classes)) + ggtitle(" PCA of the training and testing set \n mRNAArray Data of GBM") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") 
p1

```




