
---
title: "SBC on the GBM dataset from the whole Firebrowse Database(TCGA)."
subtitle: "Canonical Pathway Gene Sets"
author: "Camila Duitama (Based on Ashar Ahmad's code for the SBC model)"
date: "May 2019"
params:
  iter: 100
  iter.burnin: 150
  iter.thin: 5
  pval.sbc : 0.99
  output_folder : "GSVA_files/"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE,include=TRUE,collapse=TRUE,warning=FALSE,message=FALSE, fig.width = 10, fig.height = 6)
rm(list = ls(all.names = TRUE))
load("/Volumes/GoogleDrive/My Drive/Documents/GitHub/MasterThesis/R_Scripts_First_Results/GSVA.RData")
```
# Description

The Canonical Pathway Gene Sets were downloaded from [MSigDB](http://software.broadinstitute.org/gsea/msigdb/genesets.jsp?collection=CP) and a Gene Set Enrichment Score was calculated for the GSVA method. The GSVA scores were used as inputs for the SBC model.

# Loading of libraries

Libraries needed to run the code must be loaded first

```{r LibrariesLoading, eval=FALSE}
library(GSVA)
library(sva)
library(readr)
library(ggplot2)
library(matrixcalc)
library(stats)
library(Runuran)
library(truncnorm)
library(Matrix)
library(psych)
library(VGAM)
library(MixSim)
library(statmod)
library(flexclust)
library(mixAK)
library(mclust)
library(monomvn)
library(cluster)
library(flexmix)
library(survival)
library(utils)
library(rms)
library(pec)
library(ipred)
library(verification)
library(Hmisc)
library(glmpath)
library(glmnet)
library(gplots)
library(sparcl)
library(kernlab)
library(class)
library(reshape)
library(GGally)
library(XML)
library(RCurl)
library(MASS)
library(matrixcalc)
library(stats)
library(Runuran)
library(MCMCpack)
library(VGAM)
library(statmod)
library(survcomp)
library(gtools)
library(ggplot2)
library(GenOrd)
library(plyr)
library(dplyr)
library(iCluster)
library(CCA)
library(caret)
library(mcclust)
library(Biobase)
library(mixtools)
library(knitr)
library(kableExtra)
library(devtools)
library(rmarkdown)
library(GSVAdata)
library(msigdbr)


```

# Loading/Splitting training and testing sets

Data is separated into training and testing

```{r Loading files, eval=FALSE}
######## Input ###############

Clinical_TrainingSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Second_data_split/TrainingSet/Clinical_TrainingSet.csv")
mRNAArray_TrainingSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Second_data_split/TrainingSet/mRNAArray_TrainingSet.csv")
Clinical_TestSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Second_data_split/TestSet/Clinical_TestSet.csv")
mRNAArray_TestSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Second_data_split/TestSet/mRNAArray_TestSet.csv")
indexes.train<-which(is.na(Clinical_TrainingSet$days_to_death), arr.ind=TRUE)
indexes.test<-which(is.na(Clinical_TestSet$days_to_death), arr.ind=TRUE)
training_time<-Clinical_TrainingSet$days_to_death
test_time<-Clinical_TestSet$days_to_death
for (i in indexes.train){
  training_time[i]<-Clinical_TrainingSet$days_to_last_followup[i]
}
for (i in indexes.test){
  test_time[i]<-Clinical_TestSet$days_to_last_followup[i]
}
total_time<-append(training_time,test_time)

n<-length(Clinical_TrainingSet$X1)
n.new<-length(Clinical_TestSet$X1)


#### Overall Survival Time Vector (N*1)
time<-total_time[1:n]
time<-log(time)
########### Event or Not Vector (N * 1) ####
censoring<-Clinical_TrainingSet$vital_status
######### mRNA (or miRNA expression values) (N*D) ###
mRNAArray_TrainingSet$X1<-NULL
Y.pre.train<-mRNAArray_TrainingSet

#### Overall Survival Time Vector (N*1)
time.new<-total_time[-(1:n)]
time.new<-log(time.new)
########### Event or Not Vector (N * 1) ####
censoring.new<-Clinical_TestSet$vital_status
######### mRNA (or miRNA expression values) (N*D) ###
mRNAArray_TestSet$X1<-NULL
Y.pre.test<-mRNAArray_TestSet

```

```{r Number of samples}
print(paste0("Number of patients in training set: ",as.character(n)))
print(paste0("Number of patients in test set: ",as.character(n.new)))
```

# GSVA preprocessing

```{r eval=FALSE}
gs<-list()
db_genesets<-msigdbr(species="Homo sapiens",category = "C2",subcategory="CP")
for (i in unique(db_genesets$gs_name)){
  gs[i]<-list(unique(filter(db_genesets,db_genesets$gs_name==i)$gene_symbol))
}

Y.pre.train.gsva <- t(gsva(t(Y.pre.train),gs,mx.diff=TRUE,verbose=FALSE,parallel.sz=1))
Y.pre.test.gsva <- t(gsva(t(Y.pre.test), gs, mx.diff=TRUE, verbose=FALSE, parallel.sz=1))

Y.pre.train.plage <- t(gsva(t(Y.pre.train),gs,mx.diff=TRUE,verbose=FALSE,parallel.sz=1,method="plage"))
Y.pre.test.plage <-t(gsva(t(Y.pre.test), gs, mx.diff=TRUE, verbose=FALSE, parallel.sz=1,method="plage"))

Y.pre.train.zscore <- t(gsva(t(Y.pre.train),gs,mx.diff=TRUE,verbose=FALSE,parallel.sz=1,method="zscore"))
Y.pre.test.zscore <-t(gsva(t(Y.pre.test), gs, mx.diff=TRUE, verbose=FALSE, parallel.sz=1,method="zscore"))

Y.pre.train.ssgsea <- t(gsva(t(Y.pre.train),gs, method="ssgsea",mx.diff=TRUE,verbose=FALSE,parallel.sz=1))
Y.pre.test.ssgsea <-t(gsva(t(Y.pre.test), gs, method="ssgsea", mx.diff=TRUE, verbose=FALSE, parallel.sz=1))

```


## Data visualization

### Censoring data for training and test set
```{r Table for censoring information of the training set}
table(censoring) %>%  
  kable(caption = "Censoring frequency for the training set") %>%
  kable_styling(bootstrap_options = c("striped", "hover","responsive"),full_width = T, position = "center")

```

```{r Table for censoring information of the testing set}
table(censoring.new) %>%
  kable(caption = "Censoring frequency for the testing set") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")

```

### PCA of the GSE methods

```{r PCA Zscore}
totalmRNA<-rbind(Y.pre.train.zscore,Y.pre.test.zscore)
mRNA.pca<-as.data.frame(prcomp(totalmRNA,scale=TRUE, center= TRUE)$x[,1:2])
classes<-c()
for (i in 1:length(Clinical_TrainingSet$days_to_death)){
    classes<-append(classes,"Training")
  }
for (i in 1:length(Clinical_TestSet$days_to_death)){
    classes<-append(classes,"Testing")
  }

mRNA.pca$Classes<-classes
p1 <- ggplot(mRNA.pca, aes(x=PC1, y=PC2, colour=Classes)) + ggtitle(" PCA of the training and testing set \n Zscores of the GBM-TCGA Data set") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") 
p1

```
```{r PCA PLAGE}

totalmRNA<-rbind(Y.pre.train.plage,Y.pre.test.plage)
mRNA.pca<-as.data.frame(prcomp(totalmRNA,scale=FALSE, center= FALSE)$x[,1:2])
classes<-c()
for (i in 1:length(Clinical_TrainingSet$days_to_death)){
    classes<-append(classes,"Training")
  }
for (i in 1:length(Clinical_TestSet$days_to_death)){
    classes<-append(classes,"Testing")
  }

mRNA.pca$Classes<-classes
p1 <- ggplot(mRNA.pca, aes(x=PC1, y=PC2, colour=Classes)) + ggtitle(" PCA of the training and testing set \n PLAGE scores of the GBM-TCGA Data set") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") 
p1

```
```{r PCA ssGSEA}
totalmRNA<-rbind(Y.pre.train.ssgsea,Y.pre.test.ssgsea)
mRNA.pca<-as.data.frame(prcomp(totalmRNA,scale=TRUE, center= TRUE)$x[,1:2])
classes<-c()
for (i in 1:length(Clinical_TrainingSet$days_to_death)){
    classes<-append(classes,"Training")
  }
for (i in 1:length(Clinical_TestSet$days_to_death)){
    classes<-append(classes,"Testing")
  }

mRNA.pca$Classes<-classes
p1 <- ggplot(mRNA.pca, aes(x=PC1, y=PC2, colour=Classes)) + ggtitle(" PCA of the training and testing set \n ssGSEA scores of the GBM-TCGA Data set") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") 
p1

```
```{r PCA GSVA}
totalmRNA<-rbind(Y.pre.train.gsva,Y.pre.test.gsva)
mRNA.pca<-as.data.frame(prcomp(totalmRNA,scale=TRUE, center= TRUE)$x[,1:2])
classes<-c()
for (i in 1:length(Clinical_TrainingSet$days_to_death)){
    classes<-append(classes,"Training")
  }
for (i in 1:length(Clinical_TestSet$days_to_death)){
    classes<-append(classes,"Testing")
  }

mRNA.pca$Classes<-classes
p1 <- ggplot(mRNA.pca, aes(x=PC1, y=PC2, colour=Classes)) + ggtitle(" PCA of the training and testing set \n GSVA scores of the GBM-TCGA Data set") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") 
p1

```
# Signature calculation

This chunk calculates  SBC gene signature
It's based on the idea of Univariate testing of Survival Data features



```{r SBC Signature, eval=FALSE}
######## Prefiltering of the Genes ############################### ###########################
######## Using Univariate Cox's Model for Ranking of the Survival Power of the Genes ########
todo<-list(Y.pre.train.zscore,Y.pre.train.plage,Y.pre.train.ssgsea,Y.pre.train.gsva)

output<-list() 
 
for (i in (1:4)){
  input<-as.matrix(todo[[i]])
  surv.obj <- Surv(time,censoring)
  coeff.sig <- c(0)
  
  pvalue.sig <- c(0)
  
  
  calcCox = function(x){
    q1 <- unlist(summary(coxph(surv.obj ~ ., data = as.data.frame(x))))
    return(q1$logtest.pvalue)
  }
  
  
  pvalue.sig <- apply(input,2,calcCox)
  
  
  ###### Adjusting p-values for Multiple Test Correction
  pvalue.sig.adj <- p.adjust(pvalue.sig, method = "fdr")
  
  #### As the number of features are quite variable choose first a very loose cut-off 
  
  signature.loose <- colnames(input)[(pvalue.sig.adj < params$pval.sbc)] 
  
  ### Combined the P-values
  pvalue.combined <- (pvalue.sig.adj) 
  names(pvalue.combined) <- colnames(input)
  ## Sort it
  pvalue.combined.sort <- sort(pvalue.combined)
  ## Only select those genes which are loosely in the signature
  pvalue.combined.adj <- pvalue.combined.sort[names(pvalue.combined.sort) %in% signature.loose]
  
  output[i]<-list(pvalue.combined.adj)
}  
  ### Take the top 50 genes ####
  signature.sbc <- names(output[[3]][1:25])
  
```
## Selection of Gene Set Enrichment method

```{r Which method should we select?}
mean<-list()
sd<-list()
min<-list()
max<-list()
for (i in (1:4)){
  mean<-c(mean,round(mean(output[[i]][1:25]),4))
  sd<-c(sd,round(sd((output[[i]][1:25])),4))
  min<-c(min,round(min((output[[i]][1:25])),4))
  max<-c(max,round(max(output[[i]][1:25]),4))
  
}
result<-data.frame("Method"=c("Zscore","PLAGE","ssGSEA","GSVA"))
result$Mean<-mean
result$SD<-sd
result$Min<-min
result$Max<-max
result%>%
     kable(caption = "Results of the Univariate Cox Model on the 4 Pathway scoring methods") %>%
     kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = F, position = "center")%>%
     scroll_box(width = "500px", height = "200px")

```
**NOTE:**
Since none of the p values were significant. The model was not trained.


