
---
title: "SBC on the GBM dataset from the whole Firebrowse Database(TCGA)."
subtitle: "Karnofsky corrected results +  Flexmix Initialization"
author: "Camila Duitama (Based on Ashar Ahmad's code for the SBC model)"
date: "May 2019"
params:
  iter: 100
  iter.burnin: 150
  iter.thin: 5
  pval.sbc : 0.6
  output_folder : "Karnofsky_correction+Flexmix_initialization_files/"

---

# Description

After the Karnosfky Index correction, we wanted to try a different initialization method from the standard used in the SBC model. Therefore we tried a Flexmix initialization. This are the results of the SBC model applied to a Karnofsky corrected dataset and with a Flexmix initialization.

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE,eval=FALSE,include=TRUE,collapse=TRUE,warning=FALSE,message=FALSE, fig.width = 10, fig.height = 6)
rm(list = ls(all.names = TRUE))
load("/Volumes/GoogleDrive/My Drive/Documents/GitHub/MasterThesis/R_Scripts_First_Results/Karnofsky_correction+Flexmix_initialization.RData")

```
# Loading of libraries

Libraries needed to run the code must be loaded first

```{r LibrariesLoading, eval=FALSE}
library(sva)
library(readr)
library(ggplot2)
library(matrixcalc)
library(stats)
library(Runuran)
library(truncnorm)
library(Matrix)
library(psych)
library(VGAM)
library(MixSim)
library(statmod)
library(flexclust)
library(mixAK)
library(mclust)
library(monomvn)
library(cluster)
library(flexmix)
library(survival)
library(utils)
library(rms)
library(pec)
library(ipred)
library(verification)
library(Hmisc)
library(glmpath)
library(glmnet)
library(gplots)
library(sparcl)
library(kernlab)
library(class)
library(reshape)
library(GGally)
library(XML)
library(RCurl)
library(MASS)
library(matrixcalc)
library(stats)
library(Runuran)
library(MCMCpack)
library(VGAM)
library(statmod)
library(survcomp)
library(gtools)
library(ggplot2)
library(GenOrd)
library(plyr)
library(dplyr)
library(iCluster)
library(CCA)
library(caret)
library(mcclust)
library(Biobase)
library(mixtools)
library(knitr)
library(kableExtra)
library(devtools)
library(rmarkdown)


```

# Loading/Splitting training and testing sets

Data is separated into training and testing

```{r Loading files, eval=FALSE}
######## Input ###############

Clinical_TrainingSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/First_data_split/TrainingSet/Clinical_TrainingSet.csv")
mRNAArray_TrainingSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/First_data_split/TrainingSet/mRNAArray_TrainingSet.csv")
Clinical_TestSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Second_data_split/TestSet/Clinical_TestSet.csv")
mRNAArray_TestSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Second_data_split/TestSet/mRNAArray_TestSet.csv")

indexes.train<-which(is.na(Clinical_TrainingSet$days_to_death), arr.ind=TRUE)
indexes.test<-which(is.na(Clinical_TestSet$days_to_death), arr.ind=TRUE)
training_time<-Clinical_TrainingSet$days_to_death
test_time<-Clinical_TestSet$days_to_death
for (i in indexes.train){
  training_time[i]<-Clinical_TrainingSet$days_to_last_followup[i]
}
for (i in indexes.test){
  test_time[i]<-Clinical_TestSet$days_to_last_followup[i]
}
total_time<-append(training_time,test_time)

n<-length(Clinical_TrainingSet$X1)
n.new<-length(Clinical_TestSet$X1)


#### Overall Survival Time Vector (N*1)
time<-total_time[1:n]
time<-log(time)
########### Event or Not Vector (N * 1) ####
censoring<-Clinical_TrainingSet$vital_status
######### mRNA (or miRNA expression values) (N*D) ###
mRNAArray_TrainingSet$X1<-NULL
Y.pre.train<-mRNAArray_TrainingSet

#### Overall Survival Time Vector (N*1)
time.new<-total_time[-(1:n)]
time.new<-log(time.new)
########### Event or Not Vector (N * 1) ####
censoring.new<-Clinical_TestSet$vital_status
######### mRNA (or miRNA expression values) (N*D) ###
mRNAArray_TestSet$X1<-NULL
Y.pre.test<-mRNAArray_TestSet

```

```{r Number of samples}
print(paste0("Number of patients in training set: ",as.character(n)))
print(paste0("Number of patients in test set: ",as.character(n.new)))
```

##Data visualization

### Censoring data
```{r Table for censoring information of the training set}

table(censoring) %>%  
  kable(caption = "Censoring frequency for the training set") %>%
  kable_styling(bootstrap_options = c("striped", "hover","responsive"),full_width = T, position = "center")

```

```{r Table for censoring information of the testing set}
table(censoring.new) %>%
  kable(caption = "Censoring frequency for the testing set") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")

```

### Headers of mRNA matrixes
```{r Header of mRNA matrix of the training set}
  Y.pre.train[1:5,1:5] %>%
  kable(caption = "mRNA MicroArray data for the Training set " ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")
```

```{r Header of mRNA matrix of the testing set}
  Y.pre.test[1:5,1:5] %>%
  kable(caption = "mRNA MicroArray data for the Testing set " ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = F, position = "center")
```

### PCA of training and test set
```{r PCA of the whole mRNA dataset}
totalmRNA<-rbind(mRNAArray_TrainingSet,mRNAArray_TestSet)
mRNA.pca<-as.data.frame(prcomp(totalmRNA,scale=TRUE, center= TRUE)$x[,1:2])
classes<-c()
for (i in 1:length(Clinical_TrainingSet$days_to_death)){
    classes<-append(classes,"Training")
  }
for (i in 1:length(Clinical_TestSet$days_to_death)){
    classes<-append(classes,"Testing")
  }

mRNA.pca$Classes<-classes
p1 <- ggplot(mRNA.pca, aes(x=PC1, y=PC2, colour=Classes)) + ggtitle(" PCA of the training and testing set \n mRNAArray Data of GBM") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") 
p1

```


#Signature calculation

This chunk calculates  SBC gene signature
It's based on the idea of Univariate testing of Survival Data features


```{r SBC Signature, eval=FALSE}
######## Prefiltering of the Genes ############################### ###########################
######## Using Univariate Cox's Model for Ranking of the Survival Power of the Genes ########
  

  surv.obj <- Surv(time,censoring)
  coeff.sig <- c(0)
  
  pvalue.sig <- c(0)
  
  
  calcCox = function(x){
    q1 <- unlist(summary(coxph(surv.obj ~ ., data = as.data.frame(x))))
    return(q1$logtest.pvalue)
  }
  
  
  pvalue.sig <- apply(Y.pre.train,2,calcCox)
  
  
  ###### Adjusting p-values for Multiple Test Correction
  pvalue.sig.adj <- p.adjust(pvalue.sig, method = "fdr")
  
  #### As the number of features are quite variable choose first a very loose cut-off 
  
  signature.loose <- colnames(Y.pre.train)[(pvalue.sig.adj < params$pval.sbc)] 
  
  ### Combined the P-values
  pvalue.combined <- (pvalue.sig.adj) 
  names(pvalue.combined) <- colnames(Y.pre.train)
  ## Sort it
  pvalue.combined.sort <- sort(pvalue.combined)
  ## Only select those genes which are loosely in the signature
  pvalue.combined.adj <- pvalue.combined.sort[names(pvalue.combined.sort) %in% signature.loose]
  
  
  ### Take the top 50 genes ####
  signature.sbc <- names(pvalue.combined.adj[1:50])
  
```

The SBC signature on the dataset looks like this

```{r Table for the SBC signature}
  to_plot<-data.frame(Genes=signature.sbc)
   to_plot%>%
     kable(caption = "SBC Signature") %>%
     kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = F, position = "center")%>%
     scroll_box(width = "500px", height = "200px")
```

## PCA of the signature genes

```{r PCA of the signature}
mRNA.signature<-rbind(Y.pre.train[,signature.sbc],Y.pre.test[,signature.sbc])
mRNA.signature.pca<-as.data.frame(prcomp(mRNA.signature,scale=TRUE, center= TRUE)$x[,1:2])
mRNA.signature.pca$Classes<-classes
p6 <- ggplot(mRNA.signature.pca, aes(x=PC1, y=PC2, colour=Classes)) + ggtitle(" PCA of the training and testing set \n SBC Signature mRNAArray Data of GBM") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") 
p6
```

## Survival curves for both training and testing set

```{r Survival curves for both training and testing set}
surv.fit <- survfit(Surv(total_time,append(Clinical_TrainingSet$vital_status,Clinical_TestSet$vital_status))~ classes)
p3 <- ggsurv(surv.fit, plot.cens=FALSE,main = " DPMM \n Kaplan Meier curves for training and test set")+ ggplot2::guides(linetype = FALSE)
p3
```


#Training

##Preparation of the data needed for the training of the model
```{r Preparation of Data Structures to train the SBC, eval=FALSE}
########################################################################
###### We prepare the Data Structures Needed for the Running of the SBC ####
#############################################################################

######## Verhaak Signature ###############
Verhaak_gene_signature <-read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Verhaak_Data/Verhaak_GS_according_to_Training_Set_Split1.csv", 
                                  col_names = FALSE, col_types = cols(X1 = col_skip()))
signature.vk<-Verhaak_gene_signature$X2
######## Verhaak Labels #################
labels.vk<-Clinical_TrainingSet$Subtype

#######Getting signature matrix on training and testing######
Y <- Y.pre.train[,as.character(signature.sbc)]
c.true <- as.factor(labels.vk)
c.true<-unclass(c.true)

Y.new <- Y.pre.test[,as.character(signature.sbc)]
```


## Training results

```{r Training of SBC, eval=FALSE}
##############################################################################
############ This file takes in training data for Glioblastoma #################
################################################################################

###### Scaling the data #########
Y <- scale(Y, center = TRUE, scale = TRUE)

######
N <- nrow(Y)
D <- ncol(Y)
smod <-  Surv(exp(time), censoring)

##### Initial number of clusters
k =4
F =k


##################### STATE OF THE ART TECHNIQUES #################################
##################### BASIC METHODS + SOME ADVANCED METHODS ########################

source('Comparisonx.R')
Comparisonx()

source('ComparisionFLX.R')
ComparisionFLX()


############################# PARAMETERS for GIBB's SAMPLING ####
iter = params$iter
iter.burnin = params$iter.burnin
iter.thin  = params$iter.thin
Nps = as.integer(iter/ iter.thin)


######################### Initialize the Parameters ##############################
source('initialize.R')
initialize()

###################### Start with a good configuration ###########################
source('startSBC.R')
startSBC()


########### Train the Model #########################################
source('burninDPMM.R')
burninDPMM()

source('gibbsDPMM.R')
gibbsDPMM()

########## Analyze the fit ##########################################
### Good feature selection from heatmap plus cindex plus randindex
source('MCMCanalyze.R')
MCMCanalyze()
```
```{r}
print(paste0("Time (hours) necessary for burnin :",as.character(round(sum(burnintime)/3600),digits=5)))
print(paste0("Time (hours) necessary for Gibb's sampling :",as.character(round(sum(gibbstime/3600),digits=5))))
```


## Plots of results from training

```{r Survival curves from training}

###############################
### Some plots and analysis ####
#################################
#### Generating some plots and results ON Training data ###
logrank <- survdiff(smod ~ c.final)
pval<-1 - pchisq(unlist(logrank)$chisq,df =3)
surv.fit <- survfit(smod ~ c.final)
p4 <- ggsurv(surv.fit, surv.col=c("#b16ed3","#d37f6e","#90d36e","#6ec2d3"),plot.cens=FALSE,main = " DPMM \n Kaplan Meier Estimators \n Training Set")+ ggplot2::guides(linetype = FALSE)+labs(subtitle = paste("P-value:",toString(round(pval,digits = 10))))
p4

```

```{r PCA from training}
############ Generating some Plots ##########################
pc <- prcomp(Y)
pc.pred <- predict(pc,newdata = Y)
p5 <- ggplot(as.data.frame(pc.pred), aes(x=pc.pred[,1], y= pc.pred[,2], colour= as.factor(c.final))) + ggtitle(" SBC Clustering \n VerhaakCancer Data Set") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes")+scale_color_manual(values=c("#b16ed3","#d37f6e","#90d36e","#6ec2d3"))
p5
```

```{r Plot likelihood of burning}
to_plot<-as.data.frame(likli.burnin)
p7<-qplot(seq_along(to_plot$likli.burnin), to_plot$likli.burnin)
p7
```

# Testing

```{r PCA Scaled Y and Y.new}
#Defining again Y and Y.new
Y <- Y.pre.train[,as.character(signature.sbc)]
Y.new <- Y.pre.test[,as.character(signature.sbc)]

Y<-scale(Y, scale=TRUE,center= TRUE)
Y.new<-scale(Y.new,scale=TRUE,center=TRUE)

#PCA
Y.big.unscaled<-rbind(Y,Y.new)
pc.unscaled <- prcomp(Y.big.unscaled,center=TRUE,scale = TRUE)
labels<-c.sbc
for (i in 1:dim(Y.new)[1]){labels<-append(labels,5)}
p8 <- ggplot(as.data.frame(pc.unscaled$x), aes(x=pc.unscaled$x[,1], y= pc.unscaled$x[,2], colour= as.factor(labels))) + ggtitle("Scaled data") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") +scale_color_manual(values=c("#b16ed3","#d37f6e","#90d36e","#6ec2d3","#B1B4B6"))
p8
```


```{r Testing 2, eval=FALSE}
######## Predict on New Data Set  BASED ON JUST THE MOLECULAR DATA #####################################
smod.new <- Surv(time.new, censoring.new)
source('predictCLASS.R')
#What does this function do?: Outputs a matrix that has dimension number of test samples * number of MCMC samples
#The output matrix is c.matrix.new
predictCLASS(Y.new)
## Check the predicted Rand Index

#### Choose that configuration which has the highest difference in survival curves ####################
lr <- c(0)
for (j in 1:Nps){
  lr[j] <-  1 - pchisq(unlist(survdiff(smod.new ~ c.matrix.new[,j]))$chisq,df=length(table(c.matrix.new[,j]))-1)
}

significant<-c()
for (i in 1:Nps){
  if(lr[i]<0.05 & length(table(c.matrix.new[,i]))==4){
    significant<-append(significant,i)
  }
}

if (length(significant)>0){
  index<-significant[which.min(lr[significant])]
  c.sbc.new <- c.matrix.new[,index]

}else{
 print("There was no MCMC iteration that found 4 clusters and was statistically significant (p_value=0.05)")
 c.sbc.new <- c.matrix.new[,which.min(lr)]
}

###############################################################################
################################# Other ways of obtaining predictions #########
## If we fit cluster-specific models ###########################################

pre.sbc <- c(0)
for (q in unique(c.sbc.new)){
  ind <- which(c.sbc == q)
  ind.new <- which(c.sbc.new == q)

  time.tmp <- time[ind]

  Y.tmp <- Y[ind,]
  Y.tmp.new <- Y.new[ind.new,]

  reg <- cv.glmnet(x = Y.tmp, y = time.tmp, family = "gaussian")
  if (class(Y.tmp.new)=="numeric"){ 
    pre.sbc[ind.new] <- predict(object = reg, newx = t(matrix(Y.tmp.new)), s = "lambda.min")
  }else{
    pre.sbc[ind.new] <- predict(object = reg, newx = Y.tmp.new, s = "lambda.min")
  }
}
predCIndex.sbc.aft  <<- as.numeric(concordance(smod.new ~ exp(-pre.sbc))[1])


#### Use the adhoc-prediction algorithm ####################
source('predictADHOCverhaak.R')

source('predictTIME.R')
predictchineseAFTtime(Y.new)

```

## Plot results from testing

```{r Survival curves from testing}
logrank.new <- survdiff(smod.new ~ c.sbc.new)
#df= Degrees of freedom should be number of clusters-1
pval.new<-1 - pchisq(unlist(logrank.new)$chisq,df =length(logrank.new$n)-1)
surv.fit <- survfit(smod.new ~ c.sbc.new)
p5.new <- ggsurv(surv.fit, surv.col= c("#b16ed3","#d37f6e","#90d36e"),plot.cens=FALSE,main = " DPMM \n Kaplan Meier Estimators \n Test Set")+ ggplot2::guides(linetype = FALSE)+labs(subtitle = paste("P-value:",toString(round(pval.new,digits = 5))))

par(mfrow=c(2,1))
p4
p5.new

```

```{r PCA from testing}
pc <- prcomp(Y.new,center=TRUE,scale = TRUE)
pc.pred <- predict(pc,newdata = Y.new)
p4.new <- ggplot(as.data.frame(pc.pred), aes(x=pc.pred[,1], y= pc.pred[,2], colour= as.factor(c.sbc.new))) + ggtitle(" SBC Clustering \n GBM Cancer Test Data Set") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") +scale_color_manual(values=c("#b16ed3","#d37f6e","#90d36e"))
p4.new
```


```{r Consolidated clustering results training}
table(c.sbc) %>%
  kable(caption = "Clusters for the training set") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")
```

```{r Consolidated clustering results testing}
table(c.sbc.new) %>%
  kable(caption = "Clusters for the testing set") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")
```
```{r Final table of SBC vs others, echo=TRUE}
Methods<-c("Flexmix","K-Means + Penalized AFT","K-Means + Penalized Cox","AFT","Cox","SBC")
recovCIndex<-round(c(recovCIndex.flx,recovCIndex.km.paft,recovCIndex.km.pcox,recovCIndex.na.aft,recovCIndex.na.cox,max(recovCIndex.sbc)),digits = 5)
predCIndex<-c(predCIndex.flx,"NA",predCIndex.kk.pcox,predCIndex.na.aft,predCIndex.na.cox,max(predCIndex.sbc))
sbc_and_comparisons<-as.data.frame(Methods)
sbc_and_comparisons$recovCIndex<-recovCIndex
sbc_and_comparisons$predCIndex<-predCIndex
sbc_and_comparisons %>%
  kable(caption = "SBC vs other methods") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")
```

```{r Save Results, eval=FALSE, include=FALSE}
save.image("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/MasterThesis/R_Scripts_First_Results_Camila/Karnofsky_correction+Flexmix_initialization.RData")
write.csv(c.sbc,paste0(output_folder,"c_sbc.csv"))
write.csv(c.sbc.new,paste0(output_folder,"c_sbc_new.csv"))
write.csv(signature.sbc,paste0(output_folder,"sbc_signature.csv"))
```

