
---
title: "SBC on the GBM dataset from the Training set from Firebrowse Database(TCGA)"
author: "Camila Duitama (Based on Ashar Ahmad's code for the SBC model)"
date: "May 2019"
params:
  iter: 100
  iter.burnin: 150
  iter.thin: 5
  pval.sbc : 0.6
  output_folder : "Preliminary_Results_files/"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE,include=TRUE,collapse=TRUE,warning=FALSE,message=FALSE)
rm(list = ls(all.names = TRUE))
load("/Volumes/GoogleDrive/My Drive/Documents/GitHub/MasterThesis/R_Scripts_First_Results/Preliminary_Results.RData")

```
# Loading of libraries

Libraries needed to run the code must be loaded first

```{r LibrariesLoading}
library(sva)
library(readr)
library(ggplot2)
library(matrixcalc)
library(stats)
library(Runuran)
library(truncnorm)
library(Matrix)
library(psych)
library(VGAM)
library(MixSim)
library(statmod)
library(flexclust)
library(mixAK)
library(mclust)
library(monomvn)
library(cluster)
library(flexmix)
library(survival)
library(utils)
library(rms)
library(pec)
library(ipred)
library(verification)
library(Hmisc)
library(glmpath)
library(glmnet)
library(gplots)
library(sparcl)
library(kernlab)
library(class)
library(reshape)
library(GGally)
library(XML)
library(RCurl)
library(MASS)
library(matrixcalc)
library(stats)
library(Runuran)
library(MCMCpack)
library(VGAM)
library(statmod)
library(survcomp)
library(gtools)
library(ggplot2)
library(GenOrd)
library(plyr)
library(dplyr)
library(iCluster)
library(CCA)
library(caret)
library(mcclust)
library(Biobase)
library(mixtools)
library(knitr)
library(kableExtra)
library(devtools)
library(rmarkdown)

```

## Loading/Splitting training and testing sets

Data is separated into training and testing

```{r Loading files}
######## Input ###############
Clinical_TrainingSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/First_data_split/TrainingSet/Clinical_TrainingSet.csv")
mRNAArray_TrainingSet <- read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/First_data_split/TrainingSet/mRNAArray_TrainingSet.csv")
#Code to remove the indexes where the censoring info is NaN
indexes<-which(is.na(Clinical_TrainingSet$days_to_death), arr.ind=TRUE)
total_time<-Clinical_TrainingSet$days_to_death
for (i in indexes){
  total_time[i]<-Clinical_TrainingSet$days_to_last_followup[i]
}
#### Overall Survival Time Vector (N*1)
n<-round(dim(Clinical_TrainingSet)[1]/2)
time<-total_time[1:n]
time<-log(time)
########### Event or Not Vector (N * 1) ####
censoring<-Clinical_TrainingSet[1:n,]$vital_status
######### mRNA (or miRNA expression values) (N*D) ###
mRNAArray_TrainingSet$X1<-NULL
Y.pre.train<-mRNAArray_TrainingSet[1:n,]

#### Overall Survival Time Vector (N*1)
time.new<-total_time[-(1:n)]
time.new<-log(time.new)
########### Event or Not Vector (N * 1) ####
censoring.new<-Clinical_TrainingSet[-(1:n),]$vital_status
######### mRNA (or miRNA expression values) (N*D) ###
Y.pre.test<-mRNAArray_TrainingSet[-(1:n),]

```

```{r Number of samples}
print(paste0("Number of patients in training set: ",as.character(length(time))))
print(paste0("Number of patients in test set: ",length(time.new)))
```

### Data visualization

Initial visualization of censoring, survival times and mRNAArray Data of the training and testing set

```{r Table for censoring information of the training set}

table(censoring) %>%  
  kable(caption = "Censoring frequency for the training set") %>%
  kable_styling(bootstrap_options = c("striped", "hover","responsive"),full_width = T, position = "center")


```

```{r Table for censoring information of the testing set}
table(censoring.new) %>%
  kable(caption = "Censoring frequency for the testing set") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")

```


```{r Header of mRNA matrix of the training set}
  Y.pre.train[1:5,1:5] %>%
  kable(caption = "mRNA MicroArray data for the Training set " ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")
```

```{r Header of mRNA matrix of the testing set}
  Y.pre.test[1:5,1:5] %>%
  kable(caption = "mRNA MicroArray data for the Testing set " ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = F, position = "center")
```

```{r PCA of the whole mRNA dataset}
mRNA.pca<-as.data.frame(prcomp(mRNAArray_TrainingSet,scale=TRUE, center= TRUE)$x[,1:2])
classes<-c()
for (i in 1:length(Clinical_TrainingSet$days_to_death)){
  if (i<=n){
    classes<-append(classes,"Training")
  }else{
    classes<-append(classes,"Testing")
  }
}
mRNA.pca$Classes<-classes
p1 <- ggplot(mRNA.pca, aes(x=PC1, y=PC2, colour=Classes)) + ggtitle(" PCA of the training and testing set \n mRNAArray Data of GBM") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") 
p1

```


# Signature calculation

This chunk calculates  SBC gene signature
It's based on the idea of Univariate testing of Survival Data features


```{r SBC Signature}
######## Prefiltering of the Genes ############################### ###########################
######## Using Univariate Cox's Model for Ranking of the Survival Power of the Genes ########
  

  surv.obj <- Surv(time,censoring)
  coeff.sig <- c(0)
  
  pvalue.sig <- c(0)
  
  
  calcCox = function(x){
    q1 <- unlist(summary(coxph(surv.obj ~ ., data = as.data.frame(x))))
    return(q1$logtest.pvalue)
  }
  
  
  pvalue.sig <- apply(Y.pre.train,2,calcCox)
  
  
  ###### Adjusting p-values for Multiple Test Correction
  pvalue.sig.adj <- p.adjust(pvalue.sig, method = "fdr")
  
  #### As the number of features are quite variable choose first a very loose cut-off 
  
  signature.loose <- colnames(Y.pre.train)[(pvalue.sig.adj < 0.6)] 
  
  ### Combined the P-values
  pvalue.combined <- (pvalue.sig.adj) 
  names(pvalue.combined) <- colnames(Y.pre.train)
  ## Sort it
  pvalue.combined.sort <- sort(pvalue.combined)
  ## Only select those genes which are loosely in the signature
  pvalue.combined.adj <- pvalue.combined.sort[names(pvalue.combined.sort) %in% signature.loose]
  
  
  ### Take the top 50 genes ####
  signature.sbc <- names(pvalue.combined.adj[1:50])
  
```

The SBC signature on the dataset looks like this

```{r Table for the SBC signature}
  to_plot<-data.frame(Genes=signature.sbc)
   to_plot%>%
     kable(caption = "SBC Signature") %>%
     kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = F, position = "center")%>%
     scroll_box(width = "500px", height = "200px")
```

## PCA of the signature genes

```{r PCA of the signature}
mRNA.signature<-rbind(Y.pre.train[,signature.sbc],Y.pre.test[,signature.sbc])
mRNA.signature.pca<-as.data.frame(prcomp(mRNA.signature,scale=TRUE, center= TRUE)$x[,1:2])
mRNA.signature.pca$Classes<-classes
p6 <- ggplot(mRNA.signature.pca, aes(x=PC1, y=PC2, colour=Classes)) + ggtitle(" PCA of the training and testing set \n SBC Signature mRNAArray Data of GBM") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") 
p6
```

## Survival curves for both training and testing set

```{r Survival curves for both training and testing set}
surv.fit <- survfit(Surv(total_time,Clinical_TrainingSet$vital_status)~ classes)
p3 <- ggsurv(surv.fit, plot.cens=FALSE,main = " DPMM \n Kaplan Meier curves for training and test set")+ ggplot2::guides(linetype = FALSE)
p3
```


# Training

## Preparation of the data needed for the training of the model
```{r Preparation of Data Structures to train the SBC}
########################################################################
###### We prepare the Data Structures Needed for the Running of the SBC ####
#############################################################################

######## Verhaak Signature ###############
Verhaak_gene_signature <-read_csv("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Verhaak_Data/Verhaak_GS_according_to_Training_Set_Split1.csv", 
                                  col_names = FALSE, col_types = cols(X1 = col_skip()))
signature.vk<-read_delim("/Volumes/GoogleDrive/My Drive/Documents/Life Science Informatics Master/Master thesis/Data/Verhaak_Data/Verhaak_gene_signature.csv", 
    ";", escape_double = FALSE, col_types = cols_only(`Gene Symbol` = col_guess()), 
    trim_ws = TRUE)
######## Verhaak Labels #################
labels.vk<-Clinical_TrainingSet$Subtype

#######Getting signature matrix on training and testing######
Y <- data.matrix(Y.pre.train[,as.character(signature.sbc)])
Y.new <- data.matrix(Y.pre.test[,as.character(signature.sbc)])
smod.new <- Surv(time.new, censoring.new)
c.true <- as.factor(labels.vk)[1:n]
c.true <- unclass(c.true)
c.true.new<-as.factor(labels.vk)[-(1:n)]
c.true.new<-unclass(c.true.new)


```

## Training results

```{r Training of SBC, eval=FALSE}
##############################################################################
############ This file takes in training data for Glioblastoma #################
################################################################################

###### Scaling the data #########
Y <- scale(Y, center = TRUE, scale = TRUE)

######
N <- nrow(Y)
D <- ncol(Y)
smod <-  Surv(exp(time), censoring)

##### Initial number of clusters
k =4
F =k


##################### STATE OF THE ART TECHNIQUES #################################
##################### BASIC METHODS + SOME ADVANCED METHODS ########################

source('Comparisonx.R')
Comparisonx()

source('ComparisionFLX.R')
ComparisionFLX()


############################# PARAMETERS for GIBB's SAMPLING ####
iter = 100
iter.burnin = 150
iter.thin  = 5
Nps = as.integer(iter/ iter.thin)


######################### Initialize the Parameters ##############################
source('initialize.R')
initialize()

###################### Start with a good configuration ###########################
#source('startSBC.R')
#startSBC()


########### Train the Model #########################################
source('burninDPMM.R')
burninDPMM()

source('gibbsDPMM.R')
gibbsDPMM()

########## Analyze the fit ##########################################
### Good feature selection from heatmap plus cindex plus randindex
source('MCMCanalyze.R')
MCMCanalyze()
```



## Plots of results from training

```{r Survival curves from training}

###############################
### Some plots and analysis ####
#################################
#### Generating some plots and results ON Training data ###
logrank <- survdiff(smod ~ c.final)
pval<-1 - pchisq(unlist(logrank)$chisq,df =3)
surv.fit <- survfit(smod ~ c.final)
p4 <- ggsurv(surv.fit,surv.col=c("#b16ed3","#d37f6e","#90d36e","#6ec2d3"), plot.cens=FALSE,main = " DPMM \n Kaplan Meier Estimators \n Verhaak Cancer Data Set")+ ggplot2::guides(linetype = FALSE)+labs(subtitle = paste("P-value:",toString(pval)))
p4

```

```{r PCA from training, echo=FALSE}
############ Generating some Plots ##########################
pc <- prcomp(Y)
pc.pred <- predict(pc,newdata = Y)
p5 <- ggplot(as.data.frame(pc.pred), aes(x=pc.pred[,1], y= pc.pred[,2], colour= as.factor(c.final))) + ggtitle(" SBC Clustering \n VerhaakCancer Data Set") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") +scale_color_manual(values=c("#b16ed3","#d37f6e","#90d36e","#6ec2d3"))
p5
```

```{r Plot likelihood of burning, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
to_plot<-as.data.frame(likli.burnin)
p7<-qplot(seq_along(to_plot$likli.burnin), to_plot$likli.burnin)
p7
```


# Testing

```{r }
#Defining again Y and Y.new
Y <- Y.pre.train[,as.character(signature.sbc)]
Y.new <- Y.pre.test[,as.character(signature.sbc)]

Y<-scale(Y, scale=TRUE,center= TRUE)
Y.new<-scale(Y.new,scale=TRUE,center=TRUE)

#PCA
Y.big.unscaled<-rbind(Y,Y.new)
pc.unscaled <- prcomp(Y.big.unscaled,center=TRUE,scale = TRUE)
labels<-c.sbc
for (i in 1:dim(Y.new)[1]){labels<-append(labels,5)}
p8 <- ggplot(as.data.frame(pc.unscaled$x), aes(x=pc.unscaled$x[,1], y= pc.unscaled$x[,2], colour= as.factor(labels))) + ggtitle("Unscaled data") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") +scale_color_manual(values=c("#b16ed3","#d37f6e","#90d36e","#6ec2d3","#A5A4A4"))
p8
```


```{r Testing, eval=FALSE}
smod.new <- Surv(time.new, censoring.new)
######## Predict on New Data Set  BASED ON JUST THE MOLECULAR DATA #####################################
source('predictCLASS.R')
#What does this function do?: Outputs a matrix that has dimension number of test samples * number of MCMC samples
#The output matrix is c.matrix.new
predictCLASS(Y.new)
## Check the predicted Rand Index

#### Choose that configuration which has the highest difference in survival curves ####################
lr <- c(0)
for (j in 1:Nps){
  lr[j] <-  1 - pchisq(unlist(survdiff(smod.new ~ c.matrix.new[,j]))$chisq,df=length(table(c.matrix.new[,j]))-1)
}
c.sbc.new <- c.matrix.new[,which.min(lr)]

###############################################################################
################################# Other ways of obtaining predictions #########
## If we fit cluster-specific models ###########################################

pre.sbc <- c(0)
for ( q in 1:F){
  ind <- which(c.sbc == q)
  ind.new <- which(c.sbc.new == q)

  time.tmp <- time[ind]

  Y.tmp <- Y[ind,]
  Y.tmp.new <- Y.new[ind.new,]

  reg <- cv.glmnet(x = Y.tmp, y = time.tmp, family = "gaussian")
  pre.sbc[ind.new] <- predict(object = reg, newx = Y.tmp.new, s = "lambda.min")
}
predCIndex.sbc.aft  <<- as.numeric(concordance(smod.new ~ exp(-pre.sbc))[1])


#### Use the adhoc-prediction algorithm ####################
source('predictADHOCverhaak.R')

source('predictTIME.R')
predictchineseAFTtime(Y.new)

```

## Plot results from testing

```{r Survival curves from testing}
logrank.new <- survdiff(smod.new ~ c.sbc.new)
#df= Degrees of freedom should be number of clusters-1
pval.new<-1 - pchisq(unlist(logrank.new)$chisq,df =length(logrank.new$n)-1)
surv.fit <- survfit(smod.new ~ c.sbc.new)
p5.new <- ggsurv(surv.fit,surv.col=c("#b16ed3","#d37f6e","#90d36e","#6ec2d3","#FFC300","#0F6DA2"), plot.cens=FALSE,main = " DPMM \n Kaplan Meier Estimators \n GBM Cancer Data Set")+ ggplot2::guides(linetype = FALSE)+labs(subtitle = paste("P-value:",toString(round(pval.new,digits = 5))))
p5.new
```

```{r PCA from testing}
pc <- prcomp(Y.new)
pc.pred <- predict(pc,newdata = Y.new)
p4.new <- ggplot(as.data.frame(pc.pred), aes(x=pc.pred[,1], y= pc.pred[,2], colour= as.factor(c.sbc.new))) + ggtitle(" SBC Clustering \n GBM Cancer Test Data Set") + geom_point(shape=19) + labs(y = "PC1", x = "PC2", colour = "Classes") +scale_color_manual(values=c("#b16ed3","#d37f6e","#90d36e","#6ec2d3","#FFC300","#0F6DA2"))
p4.new
```

```{r Consolidated clustering results, echo=FALSE}
table(c.sbc) %>%
  kable(caption = "Clusters for the training set") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")
```

```{r}
table(c.sbc.new) %>%
  kable(caption = "Clusters for the testing set") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")
```
```{r Final table of SBC vs others}
Methods<-c("Flexmix","Flexmix + AFT","K-Means + Penalized AFT","K-Means + Penalized Cox","AFT","Cox","SBC + AFT","SBC")
recovCIndex<-c(recovCIndex.flx,recovCIndex.flx.aft,recovCIndex.km.paft,recovCIndex.km.pcox,recovCIndex.na.aft,recovCIndex.na.cox,max(recovCIndex.sbc.paft),max(recovCIndex.sbc))
predCIndex<-c(predCIndex.flx,"NA","NA",predCIndex.kk.pcox,predCIndex.na.aft,predCIndex.na.cox,predCIndex.sbc.aft,max(predCIndex.sbc))
sbc_and_comparisons<-as.data.frame(Methods)
sbc_and_comparisons$recovCIndex<-recovCIndex
sbc_and_comparisons$predCIndex<-predCIndex
sbc_and_comparisons %>%
  kable(caption = "SBC vs other methods") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed","responsive"),full_width = T, position = "center")
```

